---
title: "Data Manipulation and EDA"
author: "Richard G. Gardiner"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data and Packages

```{r, packages, message=FALSE}
library(tidyverse)
library(readxl)
library(grid)
library(gridExtra)
library(lme4)

theme_set(theme_light())
```

This section reads in the relevant data I have collected as well as joins some additional datasets to get political culture and and legislative control.
```{r, reading_in_data, message = FALSE}
gavelFull <- read_excel("Gavel to Gavel dataset.xlsx")
politicalCulture <- read_csv("political culture.csv")
legislativeControl <- read_excel("Legislative Control.xlsx")

```

Now I join the full gavel-to-gavel dataset with the political culture and test to make sure that there is no problem of "full rank"
```{r}
gavelWithCulture <- left_join(gavelFull, politicalCulture, 
                               by = c("Fullstate" = "State"))

table(gavelWithCulture$Moralistic, gavelWithCulture$Traditionalistic)
```

The same process, with this one being legislative control.
```{r}
gavelCultureControl <- left_join(gavelWithCulture, legislativeControl,
                                 by = c("Fullstate" = "State", "Year" = "Year"))

table(gavelCultureControl$RepublicanLegslature, gavelCultureControl$SplitLegislature)
```

Here I am changing the control variables from strings into numerics to allow for modeling.
```{r}
gavelCultureControl$Culture <- ifelse(gavelCultureControl$Moralistic == 1, "Moralistic", 
                                      ifelse(gavelCultureControl$Traditionalistic == 1, "Traditionalistic",
                                             "Individualistic"))
table(gavelCultureControl$Culture, gavelCultureControl$Culture)

gavelCultureControl$legislativeControl <- ifelse(gavelCultureControl$SplitLegislature == 1, "0", 
                                                 ifelse(gavelCultureControl$SplitLegislature == 1, "Split",
                                                        "Democratic"))

gavelCultureControl$unified <- ifelse(gavelCultureControl$SplitLegislature == 0, 1, 0)
table(gavelCultureControl$legislativeControl, gavelCultureControl$unified)
```

# Initial Analysis

First graph.  This analyses each row and sees how many rows have curbs compared to legislation that does not curb courts.  While we do see a little variation between the different cultures, my guess is that almost all of the variation occurs because of raw numbers.
```{r, curbs_by_culture}
ggplot(data = subset(gavelCultureControl, !is.na(gavelCultureControl$Culture))) +
  geom_bar(aes(x = curbing)) +
  facet_wrap(~Culture)


table(gavelCultureControl$curbing)
```

As with the graph, we see that Traditionalistic states have slightly higher rates of curbing than Moralistic and Individualistic is the lowest. Still not convinced that there is a real relationship occuring.
```{r}
gavelCultureControl %>%
  group_by(Culture) %>%
  filter(!is.na(Culture)) %>%
  filter(!is.na(curbing)) %>%
  summarise(mean = mean(curbing))
```

Trying to see bill outcomes for all legislation. 
```{r}
table(gavelCultureControl$legislativeControl, gavelCultureControl$RichardLastAction)
```

Running the same code, but creating a new object in which only curbing legislation is included.  Republican legislatures have a significant more in every category.  There are certainly more instances of Republican control, but this one may be significant enough.  Interestingly, though, is that Democratic legislatures have more enacted legislation.
```{r}
gavelCultureControlOnlyCurbs <- gavelCultureControl %>%
  filter(curbing == 1) 
  
table(gavelCultureControlOnlyCurbs$legislativeControl, gavelCultureControlOnlyCurbs$RichardLastAction)
```

The last outcome did make me want to look at years and how many instances Democrats controlled the state legislature compared to Republicans.  Republicans are in power more often, but a rather small gap compared to the amount of legislation.  Results: Republican is 241 instances, Democratic is 221, and Split is 76
```{r}
table(legislativeControl$RepublicanLegslature, legislativeControl$SplitLegislature)
```

Drilling down into the data, I now look at proportions for legislation that is introduced and then enacted:
```{r}
# republican controlled legislatures introduced court curbing bills 16.7 times a year, 18.6 for democratically controlled, and split had 11.97 
introduced <- c(3702/221, 4483/241, 910/76)
introduced 

# enacted legislation proportion
enacted <- c(197/3702, 185/4483, 46/910)
enacted
```

## Calculating Distance From Each Branch

As Dr. Steigerwalt suggested, rather than making a complex ideological distance measure with all three branches, I am creating a separate variable for ideological distance for the court and each actor (Governor, lower chamber, upper chamber).  These are simply taking the absolute difference between the branches then graphing to get a better idea.
```{r}
gavelCultureControl$court_gov_dist <- abs(gavelCultureControl$judicial - gavelCultureControl$governor)
gavelCultureControl$court_lower_dist <- abs(gavelCultureControl$judicial - gavelCultureControl$lower)
gavelCultureControl$court_upper_dist <- abs(gavelCultureControl$judicial - gavelCultureControl$upper)
```

The graphs below (which are eventually put into one grid) show the ideological distance for each relationship broken down by selection
```{r, ideological_distance_graphs, collapse = TRUE}
a <- ggplot(data = gavelCultureControl) +
  geom_boxplot(aes(x = MoreDefined, y = court_gov_dist)) +
  xlab(NULL) +
  ylab("Court - Governor")
b <- ggplot(data = gavelCultureControl) +
  geom_boxplot(aes(x = MoreDefined, y = court_lower_dist)) +
  xlab(NULL) +
  ylab("Court - Lower")
c <- ggplot(data = gavelCultureControl) +
  geom_boxplot(aes(x = MoreDefined, y = court_upper_dist)) +
  xlab("Selection System") +
  ylab("Court - Upper")

grid.arrange(a,b,c)
```


## Creating objects with time and average ideological distance


```{r, ideological_distance_over_time, collapse = TRUE}
ideology_gov_year <- gavelCultureControl %>% 
  group_by(MoreDefined, Year) %>%
  filter(!is.na(court_gov_dist)) %>%
  summarise(govMean = mean(court_gov_dist)) %>%
  filter(!(Year %in% c("2007", "43147", "2013", "2014", "2015", "2016", "2017", "2018"))) # filtering out bad data


ideology_upper_year <- gavelCultureControl %>% 
  group_by(MoreDefined, Year) %>%
  filter(!is.na(court_upper_dist)) %>%
  summarise(upperMean = mean(court_upper_dist)) %>%
  filter(!(Year %in% c("2007", "43147", "2013", "2014", "2015", "2016", "2017", "2018"))) # filtering out bad data
  
ideology_lower_year <- gavelCultureControl %>% 
  group_by(MoreDefined, Year) %>%
  filter(!is.na(court_lower_dist)) %>%
  summarise(lowerMean = mean(court_lower_dist)) %>%
  filter(!(Year %in% c("2007", "43147", "2013", "2014", "2015", "2016", "2017", "2018"))) # filtering out bad data

## graphing the data
d <- ggplot(data = ideology_gov_year) +
  geom_line(aes(x = Year, y = govMean, color = MoreDefined)) +
  scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012)) +
  xlab(NULL) +
  ylab("Mean: Court - Governor") +
  labs(colour = "Selection")

e <- ggplot(data = ideology_upper_year) +
  geom_line(aes(x = Year, y = upperMean, color = MoreDefined)) +
  scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012)) +
  xlab(NULL) +
  ylab("Mean: Court - Upper") +
  labs(colour = "Selection")

f <- ggplot(data = ideology_lower_year) +
  geom_line(aes(x = Year, y = lowerMean, color = MoreDefined)) +
  scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012)) +
  xlab(NULL) +
  ylab("Mean: Court - Lower") +
  labs(colour = "Selection")

# retention continues to be at teh high end of the list
grid.arrange(e,f)
```


Steigerwalt didn't really like the graphs, so I am trying to figure out the next graph I want to make.  I think I will be a scatterplot of ideological distance (x) and the number of curbs (y), with different colors by Elected.

```{r}
gavelCultureControlOnlyCurbs %>%
  mutate(court_lower = abs(judicial - lower),
         court_upper = abs(judicial - upper)) %>%
  group_by(Fullstate, Year, Elected, court_lower, court_upper) %>%
  summarise(curbs_in_year = sum(curbing)) %>%
  ggplot(aes(x = court_lower, y = curbs_in_year, color = Elected)) +
  geom_jitter() +
  labs(x = "Distance between Court and Lower Chamber", y = "Curbs Per Year") +
  scale_color_manual(values = c("red", "black"))

# ggsave("Curbs by Ideology and Selection - lower.jpeg")

gavelCultureControlOnlyCurbs %>%
  mutate(court_lower = abs(judicial - lower),
         court_upper = abs(judicial - upper)) %>%
  group_by(Fullstate, Year, Elected, court_lower, court_upper) %>%
  summarise(curbs_in_year = sum(curbing)) %>%
  ggplot(aes(x = court_upper, y = curbs_in_year, color = Elected)) +
  geom_jitter() +
  labs(x = "Distance between Court and Upper Chamber", y = "Curbs Per Year") +
  scale_color_manual(values = c("red", "black"))

# ggsave("Curbs by Ideology and Selection - upper.jpeg")
```





### Graphing the Relationship between ideologicaly distance and court curbing

The code below creates scatter plots that show the relationship between ideological distance and court curbing legislation.  Note the difference in scales!  Unsurprisingly, there appears to be the weakest relationship between curbs and ideological distance between the Court and Governor. 
```{r}
g <- gavelCultureControl %>%
  filter(!(Year %in% c("2007", "43147", "2013", "2014", "2015", "2016", "2017", "2018"))) %>% # filtering out bad data
  group_by(Fullstate) %>%
  filter(!is.na(court_lower_dist)) %>%
  summarise(sum = sum(curbing),
            avg_ideology = mean(court_lower_dist)) %>%
  ggplot(aes(x = avg_ideology, y = sum)) +
  geom_point() +
  geom_smooth() +
  ylab(NULL) +
  xlab("Court and Lower Chamber") +
  ylab("Court Curbing Introductions")


h <- gavelCultureControl %>%
  filter(!(Year %in% c("2007", "43147", "2013", "2014", "2015", "2016", "2017", "2018"))) %>% # filtering out bad data
  group_by(Fullstate) %>%
  filter(!is.na(court_upper_dist)) %>%
  summarise(sum = sum(curbing),
            avg_ideology = mean(court_upper_dist)) %>%
  ggplot(aes(x = avg_ideology, y = sum)) +
  geom_point() +
  geom_smooth() +
  ylab("Court Curbing Introductions") +
  xlab("Court and Upper Chamber")

grid.arrange(g, h)
```

The last chart deserves further exploration.  I decided to also run the last plot, but this time only including enacted legislation.  With the note that this is limited in the number of observations, I decided to graph it.  The results still show really no relationship between ideological distance and curbs.
```{r}
gavelCultureControl %>%
  group_by(Fullstate) %>%
  filter(!is.na(court_gov_dist)) %>%
  filter(RichardLastAction == "Enacted") %>%
  summarise(sum = sum(curbing),
            avg_ideology = mean(court_gov_dist)) %>%
  ggplot() +
  geom_point(aes(x = avg_ideology, y = sum)) +
  ylab(NULL) +
  xlab("Court and Governor") +
  geom_smooth(aes(x = avg_ideology, y = sum))
```


# Model Prep

In order to run a count model, I determined that I needed to have one row per state-year combination.  This required a number of tries, but it appears to have finally worked using the following code (though see question section below for more detail).
```{r}
count_data <- gavelCultureControl %>%
  filter(!(Year %in% c("2007", "43147"))) %>% # filtering out bad data
  group_by(State, Year) %>% 
  mutate(curbs = sum(curbing)) %>%
  distinct(State, Year, .keep_all = TRUE) %>%
  arrange(State, Year) 
  
count_data
```



Now I am adding in new "elected" variable as well as making columns for each election system broken down.  This really should have happened earlier, but better late than never.
```{r}
count_data$elected <- ifelse(count_data$Elected == "Elected", 1, 0)
count_data$partisan <- ifelse(count_data$MoreDefined == "Partisan", 1, 0)
count_data$nonpartisan <- ifelse(count_data$MoreDefined == "Nonpartisan", 1, 0)
count_data$retention <- ifelse(count_data$MoreDefined == "Retention", 1, 0)
```


### Interacting Ideology and selection system

```{r}
count_data$elected_ideology_gov <- count_data$elected * count_data$court_gov_dist
count_data$elected_ideology_upper <- count_data$elected * count_data$court_upper_dist
count_data$elected_ideology_lower <- count_data$elected * count_data$court_lower_dist

count_data$partisan_ideology_gov <- count_data$partisan * count_data$court_gov_dist
count_data$partisan_ideology_upper <- count_data$partisan * count_data$court_upper_dist
count_data$partisan_ideology_lower <- count_data$partisan * count_data$court_lower_dist

count_data$nonpartisan_ideology_gov <- count_data$nonpartisan * count_data$court_gov_dist
count_data$nonpartisan_ideology_upper <- count_data$nonpartisan * count_data$court_upper_dist
count_data$nonpartisan_ideology_lower <- count_data$nonpartisan * count_data$court_lower_dist

count_data$retention_ideology_gov <- count_data$retention * count_data$court_gov_dist
count_data$retention_ideology_upper <- count_data$retention * count_data$court_upper_dist
count_data$retention_ideology_lower <- count_data$retention * count_data$court_lower_dist

summary(count_data$elected_ideology_gov)
count_data$elected
count_data$court_gov_dist
```

### Adding in Legislative Professionalization

```{r}
professionalization <- read_csv("ncsl professionalization.csv")
str(professionalization)
count_data <- left_join(count_data, professionalization, by = c("Fullstate" = "State"))

# quick vizualization to check the data
ggplot(data = count_data) +
  geom_bar(aes(Professionalization))
```

I want to run another vizualization to test the data.
```{r}
ggplot(data = count_data) +
  geom_boxplot(aes(x = Professionalization, y = curbs, colour = Professionalization), show.legend = FALSE) +
  ylab("Number of Curbs") +
  xlab("Level of Professionalization \n(Lowest to Highest)")
```

## Last Data Cleaning

After meeting with Steigerwalt, I have decided to drop all observations where the legislature has not met.  This is primarily made up of astates that meet every other year.  To do this, I added dummy observations for the few instances where there was no legislation concerning courts that year.  

```{r}
table(count_data$Year, count_data$Fullstate)
```
This table shows the states for which there are no observations that year because they did not meet. The following table shows the number of instances for each state where I had to include dummy variables.

```{r}
table(gavelFull$Description == "No Legislation", gavelFull$State)
```




# Additional EDA

This section is mainly present to get rid of the Tableau graphs


```{r}
gavelCultureControlOnlyCurbs %>%
  count(Year) %>%
  filter(Year > 2007) %>%
  mutate(Year = as.Date(as.character(Year), format = "%Y")) %>%
  ggplot(aes(x = Year, y = n)) +
  geom_line() +
  scale_x_date() +
  ylim(c(0,250)) +
  ylab("Total Number of Court Curing Legislation Introducted") +
  theme_light()

# ggsave("Curbs over time.jpeg")


gavelCultureControlOnlyCurbs %>%
  count(Type) %>%
  mutate(Type = fct_reorder(Type, n)) %>%
  ggplot(aes(Type, n)) +
  geom_bar(stat = "identity") +
  ylab("Number of Court Curbing Legislation Introduced") +
  coord_flip() +
  theme_light() 

# ggsave("curbs by type.jpeg")

gavelCultureControlOnlyCurbs %>%
  count(RichardLastAction) %>%
  mutate(RichardLastAction = fct_reorder(RichardLastAction, n)) %>%
  ggplot(aes(RichardLastAction, n)) +
  geom_bar(stat = "identity") +
  ylab("Number of Bills") +
  xlab("Final Outcome") +
  coord_flip() +
  theme_light()

# ggsave("curbs by final outcome.jpeg")

gavelCultureControlOnlyCurbs %>%
  group_by(Fullstate, MoreDefined, Year) %>%
  summarize(total_curbs = sum(curbing)) %>%
  ungroup() %>%
  group_by(MoreDefined) %>%
  summarize(mean = mean(total_curbs)) %>%
  mutate(MoreDefined = fct_reorder(MoreDefined, mean)) %>%
  ggplot(aes(MoreDefined, mean)) +
  geom_col() +
  ylab("Average Number of Bills in a Year for a State in a Given Selection System") +
  xlab("Selection System") +
  coord_flip() +
  theme_light()

ggsave("Average curbs by system.jpeg")

gavelCultureControlOnlyCurbs %>%
  group_by(Fullstate, MoreDefined, Year) %>%
  summarize(total_curbs = sum(curbing)) %>%
  filter(Year != 2007) %>%
  ungroup() %>%
  group_by(MoreDefined, Year) %>%
  summarize(mean = mean(total_curbs)) %>%
  ungroup(MoreDefined) %>%
  mutate(MoreDefined = fct_reorder(MoreDefined, mean),
          Year = as.Date(as.character(Year), format = "%Y")) %>%
  ggplot(aes(Year, mean, color = MoreDefined)) +
  geom_line() +
  ylab("Average Number of Bills for a State By Selection System") +
  xlab("Year") +
  theme_light() +
  scale_x_date() +
  ylim(c(0, 10)) +
  scale_color_discrete(name = "Selection System")

# ggsave("averages over time and selection.jpeg")

count_data %>%
  group_by(Fullstate, Professionalization, elected) %>%
  summarize(curbs = sum(curbs, na.rm = TRUE)) %>%
  ggplot(aes(Professionalization, curbs, color = as.factor(elected))) +
  geom_jitter() +
  scale_color_discrete(name = "Selection System",
                       breaks = c(0, 1),
                       labels = c("Appointed", "Elected")) +
  theme_light() +
  xlab("Level of Professionalization") +
  ylab("Number of Curbs by State")

ggsave("curbs by professionalization.jpeg")
```

```{r}
library(usmap)
library(maps)


all_states <- map_data("state")

capitals <- us.cities %>%
  filter(capital == 2) %>%
  left_join(gavelCultureControlOnlyCurbs, by = c("country.etc" = "State")) %>%
  group_by(country.etc, lat, long, MoreDefined) %>%
  count() %>%
  filter(!(country.etc %in% c("AK", "HI")))


capitals <- as_tibble(capitals)

str(capitals)

gavelCultureControlOnlyCurbs$Fullstate <- tolower(gavelCultureControlOnlyCurbs$Fullstate)


gavelCultureControlOnlyCurbs %>%
  group_by(Fullstate, MoreDefined) %>%
  count() %>%
  left_join(all_states, by = c("Fullstate" = "region")) %>%
  ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = MoreDefined),
               color = "white", alpha = 0.3) +
  geom_point(data = capitals, aes(x = long, y = lat, size = n, color = "blue"), show.legend = FALSE) +
  theme_light() +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  xlab("") +
  ylab("") +
  theme_void() +
  labs(caption = "Larger Circles Indicate More Curbs") +
  scale_fill_discrete(name = "Selection System")


# ggsave("map of curbs.jpeg")
```


```{r}
gavelCultureControlOnlyCurbs %>%
  group_by(Fullstate, MoreDefined) %>%
  count() %>%
  left_join(all_states, by = c("Fullstate" = "region")) %>%
  ggplot() +
  geom_point(data = capitals, aes(x = long, y = lat, size = n, color = MoreDefined)) +
  borders("state") +
  theme_light() +
  xlab("") +
  ylab("") +
  theme_void() +
  labs(caption = "Larger Circles Indicate More Curbs",
       color = "Selection System",
       size = "Court Curbs") +
  scale_fill_discrete(name = "Selection System")

# ggsave("simple map of curbs.jpeg")
```


```{r}
gavelCultureControl %>%
  select(State, Year, judicial, lower, upper) %>%
  filter(Year == 2011) %>%
  distinct() %>%
  select(-Year) %>%
  gather(Branch, Ideology, -State) %>%
  mutate(Branch = str_to_title(Branch)) %>%
  ggplot(aes(x = Ideology, y = State, shape = Branch, color = Branch)) +
  geom_point() +
  labs(y = "") +
  scale_color_manual(values = c("red", "black", "black"))

# ggsave("Ideology Graph.jpeg")
```




```{r}
# write_csv(count_data, "count_data.csv", na = "")
# write_csv(gavelCultureControl, "gavelCultureControl.csv", na = "")
# write_csv(gavelFull, "gavelFull.csv", na = "")
```


